Lower Bounds on the Minimax Risk for the \\ Source Localization Problem

### 2017-01-23 paper.tex
WIP: Incorporate edits from comments.

The ``source localization'' problem is one in which we wish to estimate the
location of a point source observed through a diffusive medium using an array
of sensors. The diffusive medium is modeled as a low-pass filter, and the noise
is taken to be additive, white and gaussian.

We give lower bounds on the minimax risk (mean squared-error in location) in
estimating the location of the source, which applies to all estimators, for
\textcolor{red}{compactly supported (am I using this correctly?)}
Lipschitz-continuous impulse responses of the low-pass filter, when using a
unformly distributed sensor array. We show that the minimax error decays to
zero inversely in the number of sensors. We also perform a preliminary
numerical analysis for a more physical model that views sensors as integrators,
wherein the bound saturates before going to zero for large sensor numbers. The
benefit of increased sensor density is seen only when the signal-to-noise ratio
is high. Our bounds are the first to give a scaling in terms of the number of
sensors used.

While our results are shown for sources located on a one-dimensional circular
domain, they can be easily extended to sources on one-dimensional line segments
and d-dimensional boxes. The technique for deriving these minimax bounds also
sheds light on the sensor placement problem: increased resolution in a narrow
region can be obtained by placing sensors \emph{around} the region of interest.

@@

The source localization problem is one that arises in many fields in different
forms. Our principal motivation comes from Electroencephalography (EEG), a
non-invasive brain measurement modality which uses sensors (electrodes) placed
on the scalp to sense electric potentials produced by neuronal activity within
the brain~\cite{Nunez2006Electric}. The neurons (or groups of neurons) which
produce this activity are frequently modeled as current dipoles, and it is
often of clinical or neuroscientific interest to estimate their
locations~\cite{Baillet2001Electromagnetic}. For a single activated dipole
within the brain (which is modeled as being localized to a point), the
electrodes effectively sample a diffuse (blurred) representation of the point
source, as a result of the spatial low-pass filtering effect of the different
layers between the brain surface and scalp surface (cerebrospinal fluid, skull
and skin).

Recent developments~\cite{Grover2016Information} have suggested that source
localization can benefit greatly from an increase in sensor density. This
conclusion is based primarily on a Nyquist rate analysis, and on an
information-theoretic bound (first derived in~\cite{Grover2016Fundamental}) on
the accuracy of source localization algorithms. This information-theoretic
bound, however, was derived for the unsampled, continuous-space potential on
the scalp. In other words, it was a result asymptotic in the number of sensors.
The claim in~\cite{Grover2016Information} is therefore not fully substantiated
by this analysis, since it does not provide a lower bound showing improvement
with increasing sensor density.

Earlier work by Mosher et al.~\cite{Mosher1993Error} also sought to derive
lower bounds on source localization error by deriving Cramer-Rao lower bounds.
They also give some preliminary numerical results showing improvement with
increased numbers of sensors. However, this result is also not completely
satisfactory because Cramer-Rao bounds do not apply to biased estimators.
Commonly used source localization algorithms (such as the Minimum Norm
Estimate~\cite{Hamalainen1994Interpreting}) on the other hand, are \emph{known}
to be biased~\cite{Lin2006Assessing}.

This paper seeks to address the shortcomings of earlier methods by deriving
lower bounds on the minimax risk for the source localization problem, which
apply to \emph{all} estimators (biased \emph{and} unbiased), and which can give
a scaling in terms of the number of sensors used (for a fixed sensor placement
configuration). Computing the lower bound for the ``real brain model'' (where
the complete geometry of the brain surface is accounted for) is rendered very
difficult since analytical solutions for the EEG signal, given the brain
activity, do not exist. Previous lower bounds are based on ``spherical head
models''~\cite{Nunez2006Electric,Grover2016Information} which \emph{have}
analytical solutions. But these are still not easily amenable to analytical
lower bounds because the spherical surface does not permit uniform
sampling~\cite{Heath1956Thirteen}, and because the solution is expressed in
terms of special functions. We therefore restrict our analysis to a source
localization problem on a one-dimensional domain (to be described in detail in
Section~\ref{sec:source-localization}).

The one-dimensional toy problem is an effective tool for understanding bounds
on source localization accuracy in other settings as well, such as auditory
localization of a sound using a microphone array (although typically, very
different approaches are employed there, when compared to EEG source
localization). A vast literature on linear inverse problems
(see~\cite{Bal2012Introduction} for an introduction to this field) and
deconvolution algorithms exists, and has addressed the reduced one-dimensional
problem in broad settings
(see~\cite{Cavalier2002Sharp,Efromovich1997Robust,Ibragimov1981Statistical} and
references therein). However, our setting and interpretation appears to be
unique, since most prior work focuses on recovering a whole function (given
certain smoothness constraints), rather than locating a point source. Work that
does address point sources, to our knowledge, does not address scaling in the
number of sensors, since problem settings and applications differ greatly.

Hence, we believe that this paper is the first to give lower bounds on source
localization error (measured in an $\ell_2$-distance sense) in estimating the
location of a one-dimensional point source, which is observed by sensors
through a diffusive medium (treated as a low-pass filter), and corrupted by
additive Gaussian noise. We describe the problem setting in detail in
Section~\ref{sec:source-localization}, the minimax techniques and our main
results in Section~\ref{sec:minimax-lower-bounds}. Finally, we conclude in
Section~\ref{sec:discussion} with a discussion on how our results can be
extended to other simple settings, and on implications for sensor placement.

%\textcolor{red}{Introduction needs to be re-written.}

%Electroencephalograhy (EEG) is a system used to record the electrical activity
%of the brain. It is a non-invasive system, which may use anywhere from around
%10 to 250 electrodes placed on the scalp, to sense electric potentials. These
%potentials are generated by neuronal activity within the
%brain~\cite{Buzsaki2012Origin}. These sources of neural activity -- neurons or
%groups of neurons -- are usually modeled as current
%dipoles~\cite{Nunez2006Electric}.

%Often, it is of clinical or scientific interest to reconstruct brain activity
%patterns from EEG sensor measurements. This process is what is broadly referred
%to when we speak of the term ``source localization''. In this paper, I restrict
%the discussion to that of localizing a \emph{single} dipole, and refer to this
%process as ``dipole source localization''. Several algorithms for source
%localization have been proposed over the years
%(see~\cite{Baillet2001Electromagnetic} for an excellent review). However, until
%recently, there has been only one work that has theoretically examined the
%fundamental limits of source localization accuracy.

%Mosher et.\ al.~\cite{Mosher1993Error} first propounded Cramer-Rao lower bounds
%for unbiased source localization algorithms. Their work also forms the basis
%for setting up the source localization problem and deriving more universally
%applicable bounds in this paper. The Cramer-Rao lower bound is not in itself a
%completely satisfactory answer to the question of fundamental limits for source
%localization accuracy. This is because Cramer-Rao bounds only apply to unbiased
%estimators (or to estimators of \emph{known} bias). In practice, however, many
%popular source localization algorithms have been shown to be biased, e.g.\
%the Minimum Norm Estimate (MNE)~\cite{Hamalainen1994Interpreting} is known to
%bias the solution towards the surface of the head~\cite{Lin2006Assessing}. It
%is also known that unbiased estimators can have worse mean-squared-error
%performance than biased estimators. So Cramer-Rao bounds are useful only to a
%degree, and there may be benefit in deriving bounds that hold for all
%estimators.

%To this end, Grover~\cite{Grover2016Fundamental} recently proposed a new lower
%bound for the error in localizing a single dipole. This bound used techniques
%from information theory, for bounding the source localization error by
%expressing it in terms of the mutual information across some channel, and then
%bounding the mutual information by the capacity of the channel. In deriving
%this bound, however, the scalp potential was never sampled; equivalently, it
%was assumed that an \emph{infinite} number of sensors was available to sample
%the EEG signal at every point on the scalp. Clearly, this is a poor assumption
%and could result in a very loose bound, because it allows for more information
%to be available at the receiver than otherwise possible. The lower bound also
%relies on certain relaxations that allow the dipole to distribute its power in
%non-physical ways in frequency domain. This once again makes the bound loose by
%assigning the channel a higher capacity than what it could otherwise have.

%To address the shortcomings of the Cramer-Rao and information theoretic lower
%bounds, I apply two minimax methods to the problem of computing a lower bound
%for localization error in the idealized EEG dipole source localization problem.
%I briefly review two minimax methods -- Le Cam's, and Fano's local
%method~\cite{Duchi2015Information} -- and describe my approach in deriving the
%minimax lower bound using each of these for spherical and one-dimensional head
%models.

@@

We assume that the point source is located on a circle of circumference $S$ (a
circular domain enables us to use symmetry arguments to simplify the proof). We
can view this domain as a line, on which signals are periodic with period $S$.
The point source is therefore located somewhere within one period. We denote
the set of possible locations by $\Theta = [0, S)$. Due to the periodic nature
of the domain, a source located at position $\theta \in \Theta$ implies the
presence of sources at $s = \theta + kS$,~$\forall \, k \in \mathbb Z$ (using
`$s$' for the spatial variable).

@@

Sensors are assumed to be uniformly distributed over the domain, i.e., if there
are $m$ sensors, they are placed at locations $s = 0$, $S/m$, $2S/m$,~\dots,
$(m{-}1)S/m$ (the offset of the first sensor is arbitrary, so without loss of
generality we take it to be 0). The periodicity of the space ensures that we
automatically also have sensors at $S$, $S{+}S/m$,~\dots\@ The lower bounds we
provide are for \emph{this specific sensor configuration}. For a discussion on
why this configuration might be an appropriate choice in the minimax setting,
and on non-uniform sensor placement, see section~\ref{sec:discussion}.

@@

All signals on the aforementioned circular domain are of the form
$f:\Theta\mapsto\mathbb{R}$.  The point source located at $\theta$ is
represented by the impulse signal, $f(s;\theta) = \delta(s - \theta)$, where
$\delta(\cdot)$ is the Dirac delta function.  The sensors observe this signal
through the diffusive medium, which, intuitively speaking, blurs the impulse.
More concretely, we assume that the medium is linear and shift-invariant, so
that it can be represented by a spatial impulse response (blurring would then
correspond to low-pass filtering). Let this impulse response be given by
$g(s)$. Then, the noiseless, continuous-space signal (post-filtering and
pre-sampling) is given by the convolution, $x(s; \theta) = (g*f)(s) = g(s -
\theta)$. Here, we further make the simplifying assumption that $g(s)$ has a
sufficiently restricted support, so that aliasing effects are avoided, and
$x(s; \theta)$ is always well-defined. To be precise, $g(s) = 0$ when $\abs{s}
> w/2$, where the ``width'' $w$ of the impulse response satisfies $w < S / 2$
(the reason for the factor of $1/2$ will become clear in a later section).

The sensors sample this continuous-space shifted impulse response, with some
additive noise. We denote the noiseless sampled version of $x(s; \theta)$ by
the $m$-length vector $\v x(\theta)$:
\begin{equation} \label{eq:sampled-signal}
	\v x(\theta) = \bigg[x(0; \theta), \ldots, x\Big(\frac{kS}{m}; \theta\Big), \ldots, x\Big(\frac{(m-1)S}{m}; \theta\Big)\bigg]^T
\end{equation}
where $m$ is the number of sensors and $k \in \{0, 1, \ldots, m-1\}$. The
additive noise is given by $\v \epsilon$ (to be described shortly), and the
noisy samples are represented by $\v y$:
\begin{equation} \label{eq:sensor-obs}
	\v y = \v x(\theta) + \v \epsilon
\end{equation}
The complete signal model described in this section is summarized in
Fig.~\ref{fig:signal-model}.

@@

The noise $\v \epsilon$, introduced in equation~\eqref{eq:sensor-obs}, is
the only source of uncertainty in the problem. For our purposes, we assume that
$\v \epsilon$ is zero-mean and Gaussian, $\v \epsilon \sim \mathcal{N}(\v 0, \m
\Sigma)$. Two kinds of covariance matrices are of particular interest. The
first is dubbed the ``sensor noise'' setting, wherein $\m\Sigma = \sigma^2 \m
I$, i.e. the sensors are afflicted by i.i.d.\ noise. The second is called the
``source noise'' setting, wherein $\m\Sigma$ has off-diagonal terms. \textcolor{blue}{We
restrict our attention to the sensor noise setting in this paper.}

With the addition of Gaussian noise, each possible source location $\theta$
gives rise to a different distribution at the sensors, denoted by
\begin{equation} \label{eq:p-theta}
	P(\theta) = \mathcal{N}(\v x(\theta), \m \Sigma).
\end{equation}
$\v y$ is therefore one sample from $P(\theta)$. The space of distributions
produced by all possible source locations is $\mathcal{P} = \{P(\theta) :
\theta \in \Theta \}$. We are interested in computing lower bounds for the loss
function $\Phi(\rho(\theta, \hat\theta)) = \abs{\theta - \hat\theta}^2$, where
$\hat\theta$ is some estimate of the location based on $n$ trials (i.e. $n$
realizations) of the noisy sensor observations $\v y$.

@@

\textcolor{blue}{
Several techniques for deriving lower bounds on the minimax risk are described
in~\cite{Tsybakov2009Introduction}. We follow the excellent tutorial written by
John Duchi~\cite{Duchi2015Information} to briefly outline the steps involved in
deriving these bounds.
}

@@

@@

The upper bound in~\eqref{eq:delta-bound} translates into a lower
bound for~\eqref{eq:lb-after-pinsker}:
\begin{align}
	\mathfrak{M}_n(\mathcal{P}, \Phiorho) &\geq \frac{\delta^2}{2} \Bigg[ 1 - \sqrt{\frac{n \cdot 4 \kappa^2 \delta^2}{2\sigma^2} \bigg( \frac{(w + 2\delta) m}{S} + 1\bigg)} \Bigg] \\
	&= \frac{\delta^2}{2} \Bigg[ 1 - \sqrt{\frac{2n \kappa^2 \delta^2}{\sigma^2} \bigg( \frac{(w + 2\delta) m}{S} + 1\bigg)} \Bigg] \label{eq:lb-after-delta-bound}
\end{align}
Asymptotically, we choose smaller values of $\delta$ as the number of sensors
grows large. Hence, \textcolor{blue}{neglecting edge effects} and terms of order $\delta^3$, we
tighten the bound by choosing $\delta$ as a function of the remaining variables
in order to have $\sqrt{\frac{2nm\kappa^2\delta^2 w}{\sigma^2 S}} =
\frac{1}{2}$. This is achieved for $\delta = \sqrt{\frac{\sigma^2
S}{8nm\kappa^2 w}}$, so that for $\Phi(\delta) = \delta^2$,
equation~\eqref{eq:lb-after-delta-bound} becomes
\begin{equation}
	\mathfrak{M}_n(\mathcal{P}, \Phiorho) \geq \textcolor{blue}{\frac{1}{32}} \frac{\sigma^2 S}{nm\kappa^2 w}.
\end{equation}
This completes the proof.
\end{IEEEproof}

@@

The lower bounding \emph{technique} itself gives certain inisights on the issue
of sensor placement. Suppose we wanted to achieve a higher resolution in a
restricted interval of the parameter space, i.e.\ $\theta \in [a, b] \subset
\Theta$, such that $\abs{b - a} \ll w$. Then, it might be of greatest benefit
to place sensors \emph{not} precisely over $[a, b]$, but rather in the region
\emph{surrounding} $[a, b]$, corresponding to the \emph{largest change} in the
impulse response. Indeed, it is possible to derive an approximate lower bound
using Taylor's approximation, which shows that the KL-divergence between
distributions is maximized (and hence error minimized) at regions where the
slope of $x(s;\theta)$ is largest.

@@

For the application of EEG source localization, it might be possible to find
algorithms which achieve similar scalings as the lower bounds presented here.
It is also hoped that the techniques presented here will help us derive
analytical bounds in the more difficult setting.

\section*{Acknowledgements}

Praveen Venkatesh is partially supported by the Dowd Fellowship from the
College of Engineering at Carnegie Mellon University. The authors would like to
thank Philip and Marsha Dowd for their support and encouragement.

>>>>

The ``source localization'' problem is one in which we estimate the location of
a point source observed through a diffusive medium using an array of sensors.

We give lower bounds on the minimax risk (mean squared-error in location) in
estimating the location of the source, which apply to all estimators, for
\textcolor{blue}{certain classes} of impulse responses of the low-pass filter,
when using a unformly distributed sensor array. We show that
\textcolor{blue}{for sensors of a fixed size}, the lower bound decays with
increasing numbers of sensors. We also perform a preliminary numerical analysis
to understand the effect of increasing the number of sensors to infinity by
shrinking their size, wherein the bound saturates for large sensor numbers.
\textcolor{blue}{In the second scenario, it is seen that there is greater
benefit to increasing the number of sensors} as the signal-to-noise increases.
Our bounds are the first to give a scaling for the minimax risk in terms of the
number of sensors used.

%While our results are derived for sources located on a one-dimensional circular
%domain, they can be easily extended to sources on one-dimensional line segments
%and d-dimensional boxes. The technique for deriving these minimax bounds also
%sheds light on the sensor placement problem: increased resolution in a narrow
%region can be obtained by placing sensors \emph{around} the region of interest.

@@

The source localization problem arises in many fields in different forms. Our
principal motivation comes from Electroencephalography (EEG), a non-invasive
brain measurement modality which uses electrodes placed on the scalp to sense
electric potentials produced by neuronal activity within the
brain~\cite{Nunez2006Electric}. The neurons (or groups of neurons) which
produce this activity are usually modeled as current dipoles, and it is often
of clinical or neuroscientific interest to estimate their
locations~\cite{Baillet2001Electromagnetic}. For a single activated dipole
within the brain (which is modeled as a point source), the electrodes
effectively sample a diffuse (blurred) representation of the dipole, as a
result of the spatial low-pass filtering effect of the different layers between
the brain surface and scalp surface (cerebrospinal fluid, skull and skin).

Recent developments~\cite{Grover2016Information} have suggested that source
localization can benefit greatly from an increase in sensor density. This
conclusion is based primarily on a Nyquist rate analysis, and on an
information-theoretic bound~\cite{Grover2016Fundamental} on the accuracy of
source localization algorithms. This information-theoretic bound, however, was
derived for the unsampled, continuous-space potential on the scalp, or
equivalently, in the limit of the number of sensors going to infinity. The
claim that source localization can benefit from increased sensor density is,
therefore, not fully substantiated by the information-theoretic bound, since
the latter does not provide a lower bound showing improvement with increasing
sensor density.

Earlier work by Mosher et al.~\cite{Mosher1993Error} also gave numerical lower
bounds on source localization error. However, the Cramer-Rao lower bounds
derived by them do not apply to biased estimators, including several commonly
used source localization
algorithms~\cite{Hamalainen1994Interpreting,Lin2006Assessing}.

This paper seeks to address the shortcomings of earlier methods by deriving
lower bounds on the minimax risk for the source localization problem, which
apply to \emph{all} estimators (biased \emph{and} unbiased), and which captures
how the bound scales with increasing numbers of sensors. Previous lower bounds
are based on ``spherical head
models''~\cite{Nunez2006Electric,Grover2016Information}, which admit analytical
solutions for the potential of a single dipole. But spherical models are not
easily amenable to analytical lower bounds which seek to capture scaling in the
number of sensors, because the spherical surface does not permit uniform
sampling~\cite{Heath1956Thirteen}. We therefore restrict our analysis to a
source localization problem on a one-dimensional ``circular'' domain (to be
described in detail in Section~\ref{sec:source-localization}).

The one-dimensional toy problem is an effective tool for understanding bounds
on source localization accuracy in other settings as well, \textcolor{red}{such
	as auditory localization of a sound using a microphone array (although
	typically, very different approaches are employed there, when compared to
EEG source localization) <shall I remove this?>}. A vast literature on linear
inverse problems and deconvolution algorithms exists
(see~\cite{Bal2012Introduction} for an introduction to this field), and has
addressed the reduced one-dimensional problem in broad settings
(see~\cite{Cavalier2002Sharp,Efromovich1997Robust,Ibragimov1981Statistical} and
references therein). However, our setting and interpretation appears to be
unique, since most prior work focuses on recovering a whole function (given
certain smoothness constraints), rather than locating a point
source~\cite{Cavalier2002Sharp}. \textcolor{blue}{The error metric used when
recovering the whole function is the mean-squared error in the space of
square-integrable functions, rather than the distance between the true and
estimated locations}. Work that \emph{does} address point sources and
location-based error metrics~\cite{Ibragimov1981Statistical} does not, to our
knowledge, address scaling in the number of sensors.

Hence, we believe that this paper is the first to give lower bounds on source
localization error (measured in an $\ell_2$-distance sense) in estimating the
location of a one-dimensional point source, which is observed by sensors
through a diffusive medium (treated as a low-pass filter), and corrupted by
additive white Gaussian noise. We describe the problem setting in detail in
Section~\ref{sec:source-localization} and the minimax techniques and our main
results in Section~\ref{sec:minimax-lower-bounds}. We conclude in
Section~\ref{sec:discussion} with a discussion on how our results can be
extended to other simple settings.

For the application of EEG source localization, it might be possible to find
algorithms which achieve similar scalings as the lower bounds presented here.
It is also hoped that the techniques presented here will help us derive bounds
in more difficult settings, such as spherical and realistic brain models.

@@

We assume that the point source is located on a circle of circumference $S$. We
use the variable `$s$' to represent a general point on this domain. We can also
view this domain as a line, on which signals are periodic with period $S$.  The
point source is therefore located somewhere within one period. We denote the
set of possible locations by $\Theta = [0, S)$. Due to the periodic nature of
the domain, a source located at position $\theta \in \Theta$ implies the
presence of sources at $s = \theta + kS$,~$\forall \, k \in \mathbb Z$.

@@

Sensors are assumed to be uniformly distributed over the domain, i.e., if there
are $m$ sensors, they are placed at locations $s = 0$, $S/m$, $2S/m$,~\dots,
$(m{-}1)S/m$ (the offset of the first sensor is arbitrary, so without loss of
generality we take it to be 0). The periodicity of the space ensures that we
automatically also have sensors at $S$, $S{+}S/m$,~\dots\@ The lower bounds we
provide are for \emph{this specific sensor configuration}. For a discussion on
why this configuration might be an appropriate choice in the minimax setting,
see section~\ref{sec:discussion}.

@@

All signals on the aforementioned circular domain are of the form
$f:\Theta\mapsto\mathbb{R}$.  The point source located at $\theta$ is
represented by the impulse signal, $f(s;\theta) = \delta(s - \theta)$, where
$\delta(\cdot)$ is the Dirac delta function.  The sensors observe this signal
through the diffusive medium, which, intuitively speaking, blurs the impulse.
More concretely, we assume that the medium is linear and shift-invariant, so
that it can be represented by a spatial impulse response. Then, blurring would
correspond to low-pass filtering.  Let this impulse response be given by
$g(s)$. Then, the noiseless, continuous-space signal (post-filtering and
pre-sampling) is given by the convolution, $x(s; \theta) = (g*f)(s) = g(s -
\theta)$. Here, we further make the simplifying assumption that $g(s)$ has a
sufficiently restricted support, so that aliasing effects are avoided, and
$x(s; \theta)$ is always well-defined. To be precise, $g(s) = 0$ when $\abs{s}
> w/2$, where the ``width'' $w$ of the impulse response satisfies $w < S / 2$
(the reason for the factor of $1/2$ will become clear in a later section). For
the proof, $g(s)$ is also assumed to be Lipschitz continuous with parameter
$\kappa$. This implies that $\abs{g(u) - g(v)} \leq \kappa \abs{u - v} \;
\forall \;\; u, v \in \Theta$.

The sensors sample the continuous-space shifted impulse response, with some
additive noise. We denote the noiseless sampled version of $x(s; \theta)$ by
the $m$-length vector $\v x(\theta)$:
\begin{equation} \label{eq:sampled-signal}
	\v x(\theta) = \bigg[x(0; \theta), \ldots, x\Big(\frac{kS}{m}; \theta\Big), \ldots, x\Big(\frac{(m-1)S}{m}; \theta\Big)\bigg]^T
\end{equation}
where $m$ is the number of sensors and $k \in \{0, 1, \ldots, m-1\}$. The
additive noise is given by $\v \epsilon$ (to be described shortly), and the
noisy samples are represented by $\v y$:
\begin{equation} \label{eq:sensor-obs}
	\v y = \v x(\theta) + \v \epsilon
\end{equation}
In the source localization problem, we receive $n$ trials\footnote{The word
``trials'' comes from neuroscience literature, where several trials of an
experiment are run with a single participant. The participant is given the same
stimulus in each trial, evoking the same response in the brain (usually with
some trial-to-trial variability). For our purposes, we assume that trials are
effectively i.i.d. realizations of sensor noise, with no trial-to-trial
variability.} (i.e., $n$ realizations) of $\v y$ according to
equation~\eqref{eq:sensor-obs}, for the same point source $\theta$, using which
we need to estimate the location of the source.

The complete signal model described in this section is summarized in
Fig.~\ref{fig:signal-model}.

@@

The noise $\v \epsilon$, introduced in equation~\eqref{eq:sensor-obs}, is the
only source of uncertainty in the problem. For our purposes, we assume that $\v
\epsilon$ is zero-mean and Gaussian, $\v \epsilon \sim \mathcal{N}(\v 0, \m
\Sigma)$. We restrict our analysis to the ``sensor noise'' setting, wherein
$\m\Sigma = \sigma^2 \m I$, i.e. the sensors are afflicted by i.i.d.\ noise.

With the addition of Gaussian noise, each possible source location $\theta$
gives rise to a different distribution at the sensors, denoted by $P(\theta) =
\mathcal{N}(\v x(\theta), \m \Sigma)$. $\v y$ is therefore one sample from the
distribution $P(\theta)$. The space of distributions produced by all possible
source locations is $\mathcal{P} = \{P(\theta) : \theta \in \Theta \}$.

The noise is also assumed to be i.i.d.\ across trials, i.e., the $n$
realizations of $\v y$ are independently drawn from $P(\theta)$, for the
same location parameter $\theta$.

@@

Several techniques for deriving lower bounds on the minimax risk are described
in~\cite{Tsybakov2009Introduction}. We follow the excellent tutorial of John
Duchi~\cite{Duchi2015Information} to outline the steps involved.

@@

Recall that we are interested in computing lower bounds on the minimax risk for
the source localization problem defined in
Section~\ref{sec:source-localization}. We measure \emph{risk} in terms of the
squared-distance between the true location and the estimated location. Hence,
our loss function is $\Phi(\rho(\theta, \hat\theta)) = \abs{\theta -
\hat\theta}^2$, where $\hat\theta$ is some estimate of the location based on
$n$ trials of the noisy sensor observations $\v y$.

@@

Combining equations~\eqref{eq:lb-after-pinsker} and~\eqref{eq:delta-bound},
\begin{equation} \label{eq:lb-after-delta-bound}
	\mathfrak{M}_n(\mathcal{P}, \Phiorho) \geq \frac{\delta^2}{2} \Bigg[1 - \sqrt{\frac{2n \kappa^2 \delta^2}{\sigma^2}\bigg(\!\frac{(w{+}2\delta) m}{S} + 1\!\bigg)} \Bigg].
\end{equation}
Asymptotically, we choose smaller values of $\delta$ as the number of sensors
grows large. Hence, \textcolor{blue}{neglecting edge effects} and terms of order $\delta^3$, we
tighten the bound by choosing $\delta$ as a function of the remaining variables
in order to have $\sqrt{\frac{2nm\kappa^2\delta^2 w}{\sigma^2 S}} =
\frac{1}{2}$. This is achieved for $\delta = \sqrt{\frac{\sigma^2
S}{8nm\kappa^2 w}}$, so that for $\Phi(\delta) = \delta^2$,
equation~\eqref{eq:lb-after-delta-bound} becomes
\begin{equation}
	\mathfrak{M}_n(\mathcal{P}, \Phiorho) \geq \textcolor{blue}{\frac{1}{32}} \frac{\sigma^2 S}{nm\kappa^2 w}.
\end{equation}
This completes the proof.
\end{IEEEproof}

@@

%The lower bounding \emph{technique} itself gives certain inisights on the issue
%of sensor placement. Suppose we wanted to achieve a higher resolution in a
%restricted interval of the parameter space, i.e.\ $\theta \in [a, b] \subset
%\Theta$, such that $\abs{b - a} \ll w$. Then, it might be of greatest benefit
%to place sensors \emph{not} precisely over $[a, b]$, but rather in the region
%\emph{surrounding} $[a, b]$, corresponding to the \emph{largest change} in the
%impulse response. Indeed, it is possible to derive an approximate lower bound
%using Taylor's approximation, which shows that the KL-divergence between
%distributions is maximized (and hence error minimized) at regions where the
%slope of $x(s;\theta)$ is largest.

@@
